---
title: "ffff"
author: "go"
date: "11/5/2020"
output: pdf_document
---


```{r}
dataModel <- list.files(path = "data/debris/", full.names = TRUE) %>%
  lapply(FUN = read_csv, col_types = "ififidddddcfcif") %>%
  reduce(rbind)



dataModel$Time <- data$Timestamp %>%
  parse_datetime(format = "%Y%m%d%H%M%S")
dataModel$Timestamp <- NULL

dataModel <- dataModel %>% select(-ListID,-ListName)

plastic_ordered <- dataModel %>% 
  filter(`Material Description` == "PLASTIC") %>% 
  select(ItemName, Quantity) %>% 
  group_by(ItemName) %>% 
  summarise(Total = sum(Quantity)) %>% 
  arrange(desc(Total))

dataModel$ItemName <- factor(dataModel$ItemName, plastic_ordered$ItemName)

recategorise2 <- function(x){
  out = ""
  if(x %in% c(1,4,6,22)){out = "Cigarette related waste"}
  if(x %in% c(2,3,7,9,10,17,23,11)) out = "Food related waste"
  if(x %in% c(8,14,15,16,18,19,21,20)) out = "Other"
  if(x %in% c(12,13)) out = "Plastic bags and Styrofoam packaging"
  if(x %in% c(5,23,24,25)) out = "Fragments"
  if(out == "") stop(paste("Error in recategorise:", x))
  return(out)
}
plastic_typesModel <- dataModel %>% 
  filter(`Material Description` == "PLASTIC") %>% 
  select(ItemName, ItemID) %>% 
  distinct() %>% 
  mutate(label = 1:n()) %>% 
  mutate(category = purrr::map(label, recategorise2)) %>%
  mutate(category = as_factor(as.character(category))) %>% 
  select(ItemID, category)

plasticModel <- dataModel %>% 
  filter(`Material Description` == "PLASTIC") %>% 
  full_join(plastic_types, by = "ItemID")


plasticN <- plasticModel %>% 
  mutate(month = month(Time, label = FALSE), 
         year = as.integer(year(Time))) %>% 
  filter(year > 2010) %>% 
  group_by(month, year, category) %>%
  summarise(`Total Quantity` = sum(Quantity)) %>% 
  ggplot(aes(x = month, y = `Total Quantity`, fill = category)) +
    geom_col(colour = "black", size = 0.2, position = "fill") +
    facet_wrap(~year, nrow = 2) +
    scale_fill_viridis(discrete = TRUE, option = "plasma") +
    xlab("Month") +
    ylab("Proportion of Items") +
    ggtitle("Rel. frequencies of observed plastic waste by category") +
    scale_x_continuous(breaks = 1:12) +
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank()) +
    guides(fill=guide_legend(title="Category"))
```

```{r}
library(dplyr)

df12N <- plasticN  %>%
  filter(year == 2012) %>% 
  group_by(year, month) %>%
  mutate(freq = `Total Quantity` / sum(`Total Quantity`))


df13N <- plasticN  %>%
  filter(year == 2013) %>% 
  group_by(year, month) %>%
  mutate(freq = `Total Quantity` / sum(`Total Quantity`))


df14N <- plasticN  %>%
  filter(year == 2014) %>% 
  group_by(year, month) %>%
  mutate(freq = `Total Quantity` / sum(`Total Quantity`))

df15N <- plasticN  %>%
  filter(year == 2015) %>% 
  group_by(year, month) %>%
  mutate(freq = `Total Quantity` / sum(`Total Quantity`))


df16N <- plasticN  %>%
  filter(year == 2016) %>% 
  group_by(year, month) %>%
  mutate(freq = `Total Quantity` / sum(`Total Quantity`))



df17N <- plasticN  %>%
  filter(year == 2017) %>% 
  group_by(year, month) %>%
  mutate(freq = `Total Quantity` / sum(`Total Quantity`))


df18N <- plasticN  %>%
  filter(year == 2018) %>% 
  group_by(year, month) %>%
  mutate(freq = `Total Quantity` / sum(`Total Quantity`)) 

df19N <- plasticN  %>%
  filter(year == 2019) %>% 
  group_by(year, month) %>%
  mutate(freq = `Total Quantity` / sum(`Total Quantity`)) 


dfTotN <- rbind(df12N, df13N, df14N, df15N, df16N, df17N, df18N, df19N)

# plot for observing the data
(time_plotfr2N <- ggplot(dfTotN, aes(x = year, y = freq, color=category, fill = category)) +
  geom_smooth(method="lm", level=0.95) +
  theme_bw() + 
  xlab("Years") +
  ylab("relative frequency") +
  ggtitle("portion of plastic") + 
  expand_limits(y=0) +
  scale_y_continuous() + 
  scale_x_continuous()+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(size=5, face="bold")))
```



We see here a graphical representation of the relative frequency of the 5 different categories of plastic debris over the years. The hint we are getting here is that "Cigarette realted waste, "Food related waste", "Fragments" seem to experiance some change, whereas "Other" and "Plastic bags and styrofoam packaging" seem to remain steady. We go on and create a model and test it on untrained data to see what is the actual case.

```{r}
# create train and test set
n <- nrow(dfTotN)  # Number of observations
ntrain <- round(n*0.75)  # 75% for training set
set.seed(314)    # Set seed for reproducible results
tindex <- sample(n, ntrain)   # Create a random index
train_dfTotN <- dfTotN[tindex,]   # Create training set
test_dfTotN <- dfTotN[-tindex,]

# modelling for category "Cigarette related waste"

train_Cigrel <- train_dfTotN %>% 
  filter(category=="Cigarette related waste") %>% 
  group_by(year)

test_Cigrel <- test_dfTotN %>% 
  filter(category=="Cigarette related waste") %>% 
  group_by(year)

set.seed(1234)
train_Cigrel.modelN <- lm(freq ~ year, data = train_Cigrel)
summary(train_Cigrel.modelN)

print("PREDICTION")
pred_Cigrel <- predict(train_Cigrel.modelN, test_Cigrel)
summary(pred_Cigrel)

actuals_predsCigrel <- data.frame(cbind(actuals=test_Cigrel$freq, predicteds=pred_Cigrel)) 
head(actuals_predsCigrel)
correlation_accuracy <- cor(actuals_predsCigrel)
min_max_accuracy <- mean(apply(actuals_predsCigrel, 1, min) / apply(actuals_predsCigrel, 1, max))  


correlation_accuracy

min_max_accuracy
```


```{r}
# modelling for category "Food related waste"
train_Foodrel <- train_dfTotN %>% 
  filter(category=="Food related waste") %>% 
  group_by(year)

test_Foodrel <- test_dfTotN %>% 
  filter(category=="Food related waste") %>% 
  group_by(year)

set.seed(1234)
train_Foodrel.modelN <- lm(freq ~ year, data = train_Foodrel)
summary(train_Foodrel.modelN)

print("PREDICTION")
pred_Foodrel <- predict(train_Foodrel.modelN, test_Foodrel)
summary(pred_Foodrel)

actuals_predsFoodrel <- data.frame(cbind(actuals=test_Foodrel$freq, predicteds=pred_Foodrel)) 
head(actuals_predsFoodrel)
correlation_accuracy <- cor(actuals_predsFoodrel)  
min_max_accuracy <- mean(apply(actuals_predsFoodrel, 1, min) / apply(actuals_predsFoodrel, 1, max))  

correlation_accuracy
min_max_accuracy
```


```{r}
# modelling for category "Fragments"
train_Frag <- train_dfTotN %>% 
  filter(category=="Fragments") %>% 
  group_by(year)

test_Frag <- test_dfTotN %>% 
  filter(category=="Fragments") %>% 
  group_by(year)

set.seed(1234)
train_Frag.modelN <- lm(freq ~ year, data = train_Frag)
summary(train_Frag.modelN)

print("PREDICTION")
pred_Frag <- predict(train_Frag.modelN, test_Frag)
summary(pred_Frag)

actuals_predsFrag <- data.frame(cbind(actuals=test_Frag$freq, predicteds=pred_Frag)) 
head(actuals_predsFrag)
correlation_accuracy <- cor(actuals_predsFrag)  
min_max_accuracy <- mean(apply(actuals_predsFrag, 1, min) / apply(actuals_predsFrag, 1, max))  

correlation_accuracy
min_max_accuracy
```


```{r}
# modelling for category "Other"
train_Other <- train_dfTotN %>% 
  filter(category=="Other") %>% 
  group_by(year)

test_Other <- test_dfTotN %>% 
  filter(category=="Other") %>% 
  group_by(year)

set.seed(1234)
train_Other.modelN <- lm(freq ~ year, data = train_Other)
summary(train_Other.modelN)

print("PREDICTION")
pred_Other <- predict(train_Other.modelN, test_Other)
summary(pred_Other)

actuals_predsOther <- data.frame(cbind(actuals=test_Other$freq, predicteds=pred_Other)) 
head(actuals_predsOther)
correlation_accuracy <- cor(actuals_predsOther)  # 5.31%
min_max_accuracy <- mean(apply(actuals_predsOther, 1, min) / apply(actuals_predsOther, 1, max))  

correlation_accuracy
min_max_accuracy
```


```{r}
# modelling for category "Plastic bags and Styrofoam packaging"
train_Plbag <- train_dfTotN %>% 
  filter(category=="Plastic bags and Styrofoam packaging") %>% 
  group_by(year)

test_Plbag <- test_dfTotN %>% 
  filter(category=="Plastic bags and Styrofoam packaging") %>% 
  group_by(year)

set.seed(1234)
train_Plbag.modelN <- lm(freq ~ year, data = train_Plbag)
summary(train_Plbag.modelN)

print("PREDICTION")
pred_Plbag <- predict(train_Plbag.modelN, test_Plbag)
summary(pred_Plbag)

actuals_predsPlbag <- data.frame(cbind(actuals=test_Plbag$freq, predicteds=pred_Plbag)) 
head(actuals_predsPlbag)
correlation_accuracy <- cor(actuals_predsPlbag) 
min_max_accuracy <- mean(apply(actuals_predsPlbag, 1, min) / apply(actuals_predsPlbag, 1, max))  

correlation_accuracy
min_max_accuracy
```

```{r}
plastic_category <-c("cigarette related waste", "food related waste","Fragments", 
                      "Other","Plastic bags and Styrofoam packaging" )

slope_scores <- c(-0.063,0.038, 0.034, 0.000, 0.001)
slope_interpretation <-c("downward", "upward", "upward", "steady", "steady")
p_value<-c("<0.05","<0.05","<0.05", ">0.05",">0.05")
adjRsquared <- c(0.4416, 0.4324, 0.2177,  -0.0146, -0.01195)
corr_accuracy<-c(0.83, 0.50, 0.41, -0.36,-0.12)
min_max_Acc<-c(0.78,0.77,0.69, 0.64 ,0.61)

score_table1 <- data.frame(plastic_category, p_value,slope_scores, slope_interpretation, adjRsquared)

score_table2 <- data.frame(plastic_category, corr_accuracy, min_max_Acc)

score_table1
score_table2
```

