---
title: "Plastic Debris"
output: 
  html_document: 
    out.width: "100%"
    fig.align: "center"
    code_folding: "hide"
author: Stuart Watt
---

# Introduction and Setup
I'm using the following packages:
```{r Setup, message=FALSE}
library(tidyverse)
library(purrr)
library(magrittr)
library(treemap)
library(mapdata)
library(viridis)
library(lubridate)
```
Logged marine debris is available for download [here](http://marinedebris.engr.uga.edu/newmap/). I'm importing data from 2010 till Feb 19th 2020. There doesn't seem to be data before 2010. The data is reported marine debris.
```{r DataImport}
data <- list.files(path = "data/", full.names = TRUE) %>% 
  lapply(FUN = read_csv, col_types = "ififidddddcfcif") %>% 
  reduce(rbind)
```
I'm going to replace the column for time as a date data type, renaming it as simply "Time": 
```{r DateParsing}
data$Time <- data$Timestamp %>% 
  parse_datetime(format = "%Y%m%d%H%M%S")
data$Timestamp <- NULL
```

# Wrangling
A quick look at the data:
```{r}
data
```
Let's first check for missing values:
```{r MissingValues}
data %>% select_if(function(x) any(is.na(x))) %>% colnames()
```
So only these columns contain missing values. We will use an explicit missing value for the location factor:
```{r}
data$Location <- data$Location %>% fct_explicit_na()
```
Lets see the amount of unique values for each column:
```{r UniqueValueCount}
data %>% apply(2, function(x) length(unique(x)))
```
Both "ListID" and "ListName" don't give us any information, so we will remove them both.
```{r}
data <- data %>% select(-ListID,-ListName)
```
Lets see if there are any "ItemNames" associated with more than one "Material Descriptions".
```{r}
data %>% select(`Material Description`, ItemName) %>%
  distinct() %$% 
  table(ItemName) %>% 
  as_tibble() %>% 
  filter(n > 1)
```
So rubber gloves are associated with two material descriptions, but otherwise a one to many relationship exists between "Material Description" and "ItemName".
```{r}
data %>% select(`Material Description`, ItemName, Quantity) %>%
  filter(ItemName == "Rubber Gloves") %>% 
  group_by(`Material Description`) %>% 
  summarise(Quantity = sum(Quantity))
```
It seems that most rubber gloves are classified as plastic rather than rubber. I'm going to search for any extra descriptions given in the observations to try and gain some insight.
```{r}
data %>% select(`Material Description`, ItemName, Description) %>%
  filter(ItemName == "Rubber Gloves", !is.na(Description))
```
All instances of rubber gloves with non-missing descriptions are categorised as plastic. We also see that the descriptions suggest that the categorisation may be innaccurate: the last two instances here have "Balloon" in the extra descriptions... why aren't they categorised as such?

Distribution of observed debris:
```{r MaterialQuantities}
data %>% select(Quantity,Description,`Material Description`) %>% 
  group_by(`Material Description`) %>% 
  summarise(Quantity = sum(Quantity)) %>% 
  ggplot(aes(x = reorder(`Material Description`, Quantity), y = Quantity)) +
    geom_col() +
    ylab("Total recorded quantity") +
    xlab("Material class") +
    coord_flip()
```
So the most populated material class is Plastic. Note that this does not necessarily mean that plastic is the largest quantity of debris, just that the individual number of items categorised is largest.

A tree map of material quantities:
```{r TreeMap}
data %>% 
  select(`Material Description`, ItemName, Quantity) %>% 
  group_by(`Material Description`, ItemName) %>% 
  summarise(Quantity = sum(Quantity)) %>% 
  treemap(index = c("Material Description", "ItemName"),
          vSize = "Quantity", draw = TRUE)
```
Cigarettes are the most common item recorded. Perhaps some of the debris is not actually from the sea, but rather from people littering by the coastline? Does debris littered on the coastline end up in the oceans?

We have locational data, so lets check for any geographical observation bias.
```{r}
world <- map_data("world")
data %>% 
  select(Latitude, Longitude, Quantity, Location, `Material Description`) %>% 
  ggplot() +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), fill = "grey", alpha = 0.5) +
    geom_hex(aes(x = Longitude, y = Latitude), bins = 50) +
    scale_fill_viridis() +
    theme_void()
```
There seems to be a strong bias towards North America in our dataset. We will try a logarithmic plot to see things more clearly:
```{r}
data %>% 
  select(Latitude, Longitude, Quantity, Location, `Material Description`) %>% 
  ggplot() +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), fill = "grey", alpha = 0.5) +
    geom_hex(aes(x = Longitude, y = Latitude, fill = stat(log(count))), bins = 50) +
    scale_fill_viridis() +
    theme_void()
```

We need to know how reliable the location data is. I'm going to filter for "united kingdom" in the location field and plot the raw coordinates.
```{r}
data %>% 
  mutate(Location = str_to_lower(Location)) %>% 
  filter(str_detect(Location, "united kingdom")) %>% 
  select(Latitude, Longitude) %>% 
  ggplot(aes(x = Latitude, y = Longitude)) +
    geom_point(position = "jitter")
```
We have a outliers here. Maybe a difference in standards used for Longitude and Latitude? Some systems put the Latitude origin close to the UK.

# Questions
## Distribution of plastic by location.
Are the distributions of plastic fairly constant for the locations with the most observations? Let's look:
```{r}
topLocations <- data %>% 
  group_by(Location) %>% 
  summarise(sumQuantity = sum(Quantity)) %>% 
  arrange(desc(sumQuantity)) %>% 
  top_n(5,sumQuantity)

data %>% 
  filter(Location %in% topLocations$Location) %>% 
  group_by(Location, `Material Description`) %>% 
  summarise(sumQuantity = sum(Quantity)) %>% 
  arrange(desc(sumQuantity)) %>% 
  ggplot(aes(x = `Material Description`, y = sumQuantity, fill = Location)) +
    geom_col()
```
We see that the Location "unknown" has the most plastic... note that this is distinct from "(Missing)", which was our original NA values. Maybe we should merge these.

## Question: Are observed plastic item proportions time invariant?

```{r}
plastic_ordered <- data %>% 
  filter(`Material Description` == "PLASTIC") %>% 
  select(ItemName, Quantity) %>% 
  group_by(ItemName) %>% 
  summarise(Total = sum(Quantity)) %>% 
  arrange(desc(Total))

data$ItemName <- factor(data$ItemName, plastic_ordered$ItemName)

data %>% 
  filter(`Material Description` == "PLASTIC") %>% 
  mutate(month = month(Time, label = TRUE), 
         year = as.integer(year(Time))) %>% 
  filter(year > 2010) %>% 
  group_by(month, year, ItemName) %>%
  summarise(`Total Quantity` = sum(Quantity)) %>% 
  ggplot(aes(x = month, y = `Total Quantity`, fill = ItemName)) +
    geom_col(colour = "black", size = 0.2, position = "fill") +
    facet_wrap(~year, nrow = 2) +
    scale_fill_hue(l=50, c=150) +
    xlab("Month") +
    ylab("Proportion of Items")
    #scale_fill_viridis_d(option = "magma")

#ggsave("pastic_debris_plot.png", width = 60, height = 30, units = "cm")
```


# 4th July and Firework link? (Karen's Idea)
```{r}
data %>% 
  filter(`Material Description` == "PLASTIC") %$%
  unique(ItemName) %>% 
  print()
```