---
title: "Plastic Debris"
output: 
  html_document: 
    out.width: "100%"
    fig.align: "center"
    code_folding: "hide"
author: Stuart Watt
---

# Introduction and Setup
I'm using the following packages:
```{r Setup, message=FALSE}
library(tidyverse)
library(purrr)
library(magrittr)
library(treemap)
library(mapdata)
library(viridis)
```
Logged marine debris is available for download [here](http://marinedebris.engr.uga.edu/newmap/). I'm importing data from 2010 till Feb 19th 2020. There doesn't seem to be data before 2010. The data is reported marine debris.
```{r DataImport}
data <- list.files(path = "data/", full.names = TRUE) %>% 
  lapply(FUN = read_csv, col_types = "ififidddddcfcif") %>% 
  reduce(rbind)
```
A quick look at the data:
```{r}
data
```

# Exploratory Data Analysis
Let's first check for missing values:
```{r MissingValues}
data %>% select_if(function(x) any(is.na(x))) %>% colnames()
```
So only these columns contain missing values. We will use an explicit missing value for the location factor:
```{r}
data$Location <- data$Location %>% fct_explicit_na()
```
Lets see the amount of unique values for each column:
```{r UniqueValueCount}
data %>% apply(2, function(x) length(unique(x)))
```
Both "ListID" and "ListName" don't give us any information, so we will remove them both.
```{r}
data <- data %>% select(-ListID,-ListName)
```

Lets see if there are any "ItemNames" associated with more than one "Material Descriptions".
```{r}
data %>% select(`Material Description`, ItemName) %>%
  distinct() %$% 
  table(ItemName) %>% 
  as_tibble() %>% 
  filter(n > 1)
```
So rubber gloves are associated with two material descriptions, but otherwise a one to many relationship exists between "Material Description" and "ItemName".
```{r}
data %>% select(`Material Description`, ItemName, Quantity) %>%
  filter(ItemName == "Rubber Gloves") %>% 
  group_by(`Material Description`) %>% 
  summarise(Quantity = sum(Quantity))
```
It seems that most rubber gloves are classified as plastic rather than rubber.
```{r}
data %>% select(`Material Description`, ItemName, Description) %>%
  filter(ItemName == "Rubber Gloves", !is.na(Description)) %>% 
  distinct()
```
All non-missing descriptions of rubber gloves are described as plastic.

Distribution of observed debris:
```{r MaterialQuantities}
data %>% select(Quantity,Description,`Material Description`) %>% 
  group_by(`Material Description`) %>% 
  summarise(Quantity = sum(Quantity)) %>% 
  ggplot(aes(x = reorder(`Material Description`, Quantity), y = Quantity)) +
    geom_col() +
    ylab("Total recorded quantity") +
    xlab("Material class") +
    coord_flip()
```
A tree map of material quantities:
```{r TreeMap}
data %>% 
  select(`Material Description`, ItemName, Quantity) %>% 
  group_by(`Material Description`, ItemName) %>% 
  summarise(Quantity = sum(Quantity)) %>% 
  treemap(index = c("Material Description", "ItemName"),
          vSize = "Quantity", draw = TRUE)
```
Perhaps some of the debris is not actually from the sea, but rather from people littering by the coastline? Does debris littered on the coastline end up in the oceans?
```{r}
world <- map_data("world")
data %>% 
  select(Latitude, Longitude, Quantity, Location, `Material Description`) %>% 
  ggplot() +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), fill = "grey", alpha = 0.5) +
    geom_hex(aes(x = Longitude, y = Latitude), bins = 50) +
    scale_fill_viridis() +
    theme_void()
```
```{r}
data %>% mutate(Location = str_to_lower(Location)) %>% 
  filter(str_detect(Location, "united kingdom")) %>% 
  select(Latitude, Longitude) %>% 
  ggplot(aes(x = Latitude, y = Longitude)) +
    geom_point()
```
```{r}
data %>% mutate(Location = str_to_lower(Location)) %>% 
  filter(str_detect(Location, "ocean")) %>% 
  select(Location) %>% 
  distinct()
```

```{r}
topLocations <- data %>% 
  group_by(Location) %>% 
  summarise(Quantity = sum(Quantity)) %>% 
  arrange(desc(Quantity)) %>% 
  top_n(5,Quantity)

data %>% 
  filter(Location %in% topLocations$Location) %>% 
  group_by(Location, `Material Description`) %>% 
  summarise(Quantity = sum(Quantity)) %>% 
  arrange(desc(Quantity)) %>% 
  ggplot(aes(x = `Material Description`, y = Quantity, fill = Location)) +
    geom_col()
```